A Switching Planner for Combined Task and Observation Planning

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
New controller spec

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
applied at state $\state$, then the resultant state is $ ( \state \cup
\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
propositions from $\add(\detAction)$ are added to $\state$, and those
from $\delete(\detAction)$ are removed from $\state$.

@@

Whereas until now we have considered a POMDP model that is factored in
terms of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs are
negative rewards.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---I.e., reaches a state that satisfies some given
Boolean condition--- with probability $1$.  More generally, a useful
formalism for representing solutions to POMDPs is the finite-state
controller (FSC). This is a three-tuple $\langle \nodes, \selection,
\transition \rangle$ where: $\node \in \nodes$ is a set of controller
states, $\selection_\node(\action) = P(\action | \node)$ gives the
probability that the controller prescribes \action\ when in state
\node, and $\transition_\node(\action, \observation, \node') =
P(\node'|\node, \action, \observation)$ gives the probability the
controller transitions to state $\node'$, supposing we execute
\action\ in state \node\ receiving observation \observation. When
acting for finite $N$ steps, the value of a controller corresponds to
the expected cumulative reward:

\begin{equation}\label{eq:expectedvalue}
V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
\end{equation}

>>>>

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
applied at state $\state$, then the resultant state is $ ( \state \cup
\add(\detAction) ) \backslash \delete(\detAction)$ -- I.e.,
propositions from $\add(\detAction)$ are added to $\state$, and those
from $\delete(\detAction)$ are removed from $\state$.

@@

Whereas until now we have considered a POMDP model that is factored in
terms of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---I.e., reaches a state that satisfies some given
Boolean condition--- with probability $1$.  More generally, a useful
formalism for representing solutions to POMDPs is the finite-state
controller (FSC). This is a three-tuple $\langle \nodes, \selection,
\transition \rangle$ where: $\node \in \nodes$ is a set of controller
states, $\selection_\node(\action) = P(\action | \node)$ gives the
probability that the controller prescribes \action\ when in state
\node, and $\transition_\node(\action, \observation, \node') =
P(\node'|\node, \action, \observation)$ gives the probability the
controller transitions to state $\node'$, supposing we execute
\action\ in state \node\ receiving observation \observation. When
acting for finite $N$ steps from controller state $n$, the value of a
controller corresponds to the expected cumulative reward:

\begin{equation}\label{eq:expectedvalue}
V_{\pp{FSC}}(\bstate, n) = \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, n_0 = n, \bstate_0 = \bstate \bigg{]}
\end{equation}
